**The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models**

### 2021-02-04 Manuscripts/Manuscript_AM_v2.Rmd
Need to be able to put the two graphs as a single one

When the data does not follow a linear trend, the fit that an rm-ANOVA or LMEM produces may not accurately represent the trends in the data. To demonstrate this inaccurate representation, longitudinal data where a normally distributed response of two groups of 10 subjects each is simulated, and a rm-ANOVA model  (Equation \@ref(eq:linear-model)) is fitted to the data using $\textsf{R}$[@r] and the package _nlme_[@nlme] (code in the Appendix). Briefly, the mean responses for each group are linear functions with different intercepts.

@@

  return((p1 + p3) / (p2 + p4))
  # return((p1 + p3 + plot_layout(guides = "collect")) / (p2 + p4 + plot_layout(guides = "collect")))

@@

```{r, linear-model, fig.width=16, fig.height=12,  echo=FALSE,message=FALSE,fig.show='hold',fig.cap="Simulated linear responses from two groups with a rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response"}
# linear response, correlated errors (subject effect)
txt<-18
plot_example(example(fun_type = "linear", error_type = "correlated")) +
  plot_annotation(tag_levels='A',
                  title = 'Simulated with linear function and correlated errors') &
   theme(plot.title = element_text(size = 20),text=element_text(size=txt)
)
```

@@

```{r quadratic-response, fig.width=16, fig.height=16,echo=FALSE,fig.cap='Simulated quadratic responses from two groups with an rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response', message=FALSE}
# quadratic response, correlated errors (subject effect)
txt<-18
plot_example(example(fun_type = "linear", error_type = "correlated")) +
  plot_annotation(
    tag_levels='A',
  title = 'Simulated with linear function and correlated errors') &
   theme(plot.title = element_text(size = 20),text=element_text(size=txt)
)
  # theme(plot.title = element_text(size = 20))
```

>>>>

When the data does not follow a linear trend, the fit that an rm-ANOVA or LMEM produces may not accurately represent the trends in the data. To demonstrate this inaccurate representation, longitudinal data where a normally distributed response of two groups of 10 subjects each is simulated, and a rm-ANOVA model  (Equation \@ref(eq:linear-model)) is fitted to the data using $\textsf{R}$[@r] and the package _nlme_[@nlme] (code in the Appendix). Briefly, two cases for the mean responses for each group are considered: in the first case, the mean response in each group is a linear function with different intercepts and slopes (negative slope for Group 1 and positive slope for Group 2 ). In the second case, a quadratic function is used for the mean response per group (concave down for Group 1 and concave up for Group 2 ). From the mean response per group in the linear or quadratic case, individual responses for each subject in each group are simulated using covariance matrices for with compound symmetry (constant variance across time) or independent errors (errors that are generated from a a normal distribution but are not constant over time). Thus, the response per subject corresponds to the mean response per group plus the error (compound symmetry or independent). To this data, an rm-ANOVA model identical as Equation \@ref:eqn(linear-model) is fitted. The resulting simulated data, simulated errors, individual response and fitted parameters from the model are plotted.

@@

  return((p1+p3+p2+p4)+plot_layout(nrow=1))
  # return((p1 + p3 + plot_layout(guides = "collect")) / (p2 + p4 + plot_layout(guides = "collect")))

@@

```{r, linear-model-correlated, fig.width=16, fig.height=12,  echo=FALSE,message=FALSE,fig.show='hold',fig.cap="Simulated linear responses from two groups with a rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response"}
# linear response, correlated errors (subject effect)
txt<-18
plot_example(example(fun_type = "linear", error_type = "correlated")) +
  plot_annotation(tag_levels='A',
                  title = 'Simulated with linear function and correlated errors') &
   theme(plot.title = element_text(size = 20),text=element_text(size=txt)
)
```

```{r linear-model-independent, fig.width=16, fig.height=16,echo=FALSE,fig.cap='Simulated quadratic responses from two groups with an rm-ANOVA model fitted. Top row, simulated data, lines represent mean response. Bottom row, fitted model, thick lines represent predicted mean response', message=FALSE}
# quadratic response, correlated errors (subject effect)
txt<-18
plot_example(example(fun_type = "linear", error_type = "independent")) +
  plot_annotation(
    tag_levels='A',
  title = 'Simulated with linear function and independent errors') &
   theme(plot.title = element_text(size = 20),text=element_text(size=txt)
)
  # theme(plot.title = element_text(size = 20))
```

@@
